The Sub-Millimetre User Reduction Facility: an iterative map-maker for SCUBA-2

### 2012-01-21 smurf_paper.tex
write text for point sources, add scripts/figures for cosmology reduction

@@

\subsection{Point source}
In this example we produce a map of Uranus, which is a nearly
point-like source for SCUBA-2 that is commonly used as a primary flux
calibrator (observation 26 on 20111017). The first reduction uses the
defaults as described in Section~\ref{sec:algorithm}, but we manually
set the number of iterations to a fixed value of 100 to illustrate how
the solution converges in the absence of any boundary conditions.

Reduce a calibrator. Describe the use of a zero mask around the
source. Compare with \citet{wiebe2009} who used the same method to
constrain a maximum-likelihood map with poor cross-linking.

\begin{figure*}
\centering
\includegraphics[width=\linewidth]{pointmaps.pdf}
\caption{Multiple reductions of a 450\,\micron\ daisy scan of Uranus,
  all scaled between -0.002\, pW (white) and +0.002\,pW (black): (a)
  Reduction in which only common-mode subtraction is used to suppress
  low-frequency noise, and the reduction is forced to use 2 iterations
  (after the first iteration an estimate of the source flux is
  removed, and the noise is measured in the residuals to obtain
  appropriate weighting for the second and final iteration) ---
  circular streaks are caused by independent low-frequency noise that
  is not removed by the common-mode; (b) same as (a) but using 50
  iterations, illustrating that there is a degeneracy between
  large-scale structure and the common-mode (the footprint of working
  bolometers is also shown for reference); (c) reduction in which
  high-pass filtering above 0.775\,Hz (corresponding to a spatial
  scale of 200\,arcsec as indicated) is applied after common-mode
  removal, but before the map estimate --- the map-based convergence
  test is activated and the solution halts after 10 iterations, but
  leaving large-scale ringing due to the filter; (d) same as (c), but
  now the region of the map beyond the white dot-dashed circle is
  constrained to a value of zero until all but the final iteration ---
  the map is now extremely flat, there is little attenuation of the
  source flux, and the diffraction pattern is clearly seen.}
\label{fig:pointmaps}
\end{figure*}

@@

\subsection{Deep blind survey}
Heavy initial filtering, and only iteratively solve for map and
common-mode with edge constraints.

\textit{And matched filter is discussed here or superceded by
the 2D FFTs?}

>>>>

% sinc function
\def\sinc{\mathrm{sinc}}

@@

\subsection{Point source}

The accurate measurement of positions and brightnesses of point
sources are necessary in real-time to establish telescope pointing
offsets and focus. They are also necessary to measure the FCF
(absolute calibration), and hence noise performance of the instrument
in astronomical units.  In this example we produce a 450\,\micron\ map
of Uranus (observation 26 on 20111017), which is a nearly point-like
source for SCUBA-2 that is commonly used as a primary flux
calibrator. The now standard CV daisy scan pattern was used, with a
scan speed of 155 arcsec\,sec$^{-1}$. We perform several different
reductions of the data to illustrate the purpose of different model
components and the convergence properties of the solution
(Fig.~\ref{fig:pointmaps}). In all cases the maps are produced on a
grid of azimuth (horizontal) and elevation (vertical) offsets from the
position of Uranus (the origin), using 2\,arcsec pixels.

The first, simplest reduction of the data uses only the \model{COM}
model to estimate and remove the common-mode signal in order to
suppress low-frequency noise in the data. After \model{COM}, the
extinction correction is applied (\model{EXT}), and an initial map is
estimated using equal weighting for all of the detectors. This
estimate of \model{AST} is then removed from the data, and the noise
is measured in the residuals to estimate weights for the subsequent
and final iteration. The resulting map after these two iterations is
shown in Fig.~\ref{fig:pointmaps}a. While the S/N of Uranus is clearly
large ($\sim$175 as estimated by SMURF), enabling us to see the faint
sidelobes (circle and cross pattern), the map also has obvious
circular streaks and other large-scale ripples. The circular streaks
are due to the fact that \model{COM} does not account for all of the
low-frequency noise (see Fig.~\ref{fig:pspec}), and therefore each
bolometer leaves a trace of the circular scan pattern in the map as
each of their baselines slowly drift independently. Most of the
larger-scale ripples in the map, however, are probably due to
degeneracies in the map solution.

To illustrate how large-scale ripples can form (and in fact, diverge),
the same map solution is run for 50 iterations and shown in
Fig.~\ref{fig:pointmaps}b, now exhibiting a strong vertical
gradient. The degeneracy is easy to understand if the time-domain
behaviour of each model component is considered. The top panel of
Fig.~\ref{fig:degeneracy} shows the residual signals for a single
bolometer after 2 (black) and 50 (grey) iterations which are nearly
identical, yet the \emph{change} in the estimated \model{COM} (green)
and \model{AST} (red) signals between 2 and 50 iterations are
large. However, it is also clear that the estimated \model{COM} and
\model{AST} signals are \emph{complimentary}. In other words, the large
change in the \model{AST} signal is cancelled by freedom in the
\model{COM} signal to grow with opposite sign. For comparison, the
bottom panel of Fig.~\ref{fig:degeneracy} shows the pointing solution
for this section of data, and the shapes of the \model{AST} and
\model{COM} signals match the elevation component, which is aligned
with the gradient in Fig.~\ref{fig:pointmaps}b. Generically, the
calculation of \model{COM} will remove any information on angular
scales that are larger than the array footprint (outline shown in
Fig.~\ref{fig:pointmaps}b for reference), meaning that the map
solution is unconstrained on such large scales.

One simple improvement is to simply restrict the low-frequency signal
that goes into the maps using a high-pass filter. A third reduction of
the data uses the default map-making parameters, as described in
Section~\ref{sec:algorithm}, which adds the \model{FLT} model to
accomplish this task immediately prior to map estimation. We set the
filter edge based on an angular scale of 200\,arcsec, which given the
scan speed of 155\,arcsec\,sec$^{-1}$, corresponds to a frequency of
0.78\,Hz. Now using the automatic map-based convergence test (Section
???), the resulting map converges after 10 iterations
(Fig.~\ref{fig:pointmaps}c). Both the circular streaks and the
large-scale gradient are now removed, but they have been replaced by
an obvious, circularly-symmetric ringing pattern about Uranus. The
reason for this ringing is due to the fact that the hard-edged
high-pass filter in frequency space is equivalent to a $\sinc$
function-like response in map space. Since the scan pattern is
isotropic (scans equally at all position angles), and there is a
bright point-like source, the result is an azimuthally-symmetric
$\sinc$ function-like pattern in the map about Uranus. Note that due
to the iterative nature of the solution, much of the signal in the
source is actually removed for subsequent iterations of
\model{FLT}, but as in the previous case, generacies between the
map and the \model{FLT} and \model{COM} signal components can still
lead to complicated structures.

The preceding examples illustrate the need for prior constraints in
the map solution. For calibrators (and other previously known bright,
compact sources), a good, simple prior is to constrain the map to a
value of zero away from the known locations of emission. In
Fig.~\ref{fig:pointmaps}d, a map is produced in an identical manner to
the previous example, but now setting the map explicitly to zero
beyond a radius of 60\,arcsec from the location of Uranus (much larger
than the FWHM of the main lobe), for all but the final iteration. In
this case, the map converges after 22 iterations, and all of the
previous artifacts have been effectively removed. Since the map is now
flat away from the source, and constrained to a value of zero, it is
now appropriate for performing aperture photometry directly, with no
need for an additional reference annulus. This approach to constrained
map-making is similar to that employed for poorly cross-linked scans
of compact (though resolved) sources by \citet{wiebe2009} using BLAST
data.

\begin{figure*}
\centering
\includegraphics[width=\linewidth]{pointmaps.pdf}
\caption{Multiple reductions of a 450\,\micron\ daisy scan of Uranus,
  all scaled between -0.002\,pW (white) and +0.002\,pW (black): (a)
  Reduction in which only common-mode subtraction is used to suppress
  low-frequency noise, and the reduction is forced to use 2 iterations
  (after the first iteration an estimate of the source flux is
  removed, and the noise is measured in the residuals to obtain
  appropriate weighting for the second and final iteration) ---
  circular streaks are caused by independent low-frequency noise that
  is not removed by the common-mode (Uranus peak 0.14\,pW); (b) same
  as (a) but using 50 iterations, illustrating that there is a
  degeneracy between large-scale structure and the common-mode (the
  footprint of working bolometers is also shown for reference, Uranus
  peak 0.14\,pW); (c) reduction in which high-pass filtering above
  0.775\,Hz (corresponding to a spatial scale of 200\,arcsec as
  indicated) is applied after common-mode removal, but before the map
  estimate --- the map-based convergence test is activated and the
  solution halts after 10 iterations, but leaving large-scale ringing
  due to the filter (Uranus peak 0.38\,pW); (d) same as (c), but now
  the region of the map beyond the white dot-dashed circle is
  constrained to a value of zero until all but the final iteration ---
  the map is now extremely flat, there is little attenuation of the
  source flux, and the diffraction pattern is clearly seen (Uranus
  peak 0.38\,pW).}
\label{fig:pointmaps}
\end{figure*}

@@

\subsection{Deep blind survey}

\scuba\ surveys designed to detect extremely faint point-sources
(e.g. high-redshift star-forming galaxies, and debris disks) need to
obtain the best noise performance possible. The approach described
here involves three major steps: (i) generating a map that removes
most low-frequency noise sources; (ii) \emph{whitening} (essentially
flattening) the map to make the noise properties of the map easier to
deal with; and (iii) detecting point sources using a
\emph{matched-filter}. Note that variations on this general procedure
have been used extensively in the submm cosmology community using
previous instruments \citep[e.g.][]{scott2008,perera2008}.  In this
section we will reduce data of the Lockman Hole taken in 2010 during
S2SRO as a pilot project for the Cosmology Legacy Survey. It consists
of $\sim8.5$\,hours of data taken in 36 separate scans (average
15\,min. each) spread over February and March, covering an area of
about 50\,arcmin$^2$. ... full obs list ... dropped scans due to
terrible noise properties ...

The first step, map generation, is accomplished in a way similar to
the default described in the previous section (without any boundary
constraints)...

Even though the map \emph{looks} extremely flat, it is clear that
there is a mixture of faint astronomical sources, and what is probably
residual low-frequency noise causing faint ripples. With uncertainties
about what is source, and what is an artifact, the first step is to
remove the low-frequency noise in a way that does not affect
(appreciably) the real sources. To accomplish this goal, the map is
run through a whitening filter that results in a flat angular noise
power spectrum. The noise power spectrum is estimated from a
\emph{jackknife map}: maps are produced from two independent halves of
the total data set, and then a pixel-by-pixel difference is
calculated. Provided that the noise in one half of the data is
uncorrelated with that from the other half, the noise in the jackknife
map should resemble that of the total map, though larger by a factor
of $\sim\sqrt(2)$ depending on how closely the pixel variances in the
two maps are matched (if one map pixel is significantly noisier, it
will dominate the noise in the jackknife).  The astronomical signal,
however, should be cleanly removed (provided that there are no strong
time-varying signals, and also assuming that errors due to calibration
between the two halves are insignificant). The approach we have taken
to minimize systematics is to produce two halves of the jackknife
using odd and even scan numbers (i.e., each map will contain an even
sampling of data across the full dataset).

\begin{figure}
\centering
\includegraphics[width=\linewidth]{lockman_rawmaps.pdf}
\caption{Procedure for producing maps of deep fields.}
\label{fig:lockman_rawmaps}
\end{figure}

Since the \scuba\ scan strategy is usually isotropic (all position
angles scanned with roughly equal weights), we make the simplifying
assumption that the angular noise power spectrum is azimuthally
symmetric (inspection of the 2-dimensional FFT confirms
this). ... calculate filter as slightly-smoothed complement of
azavpspec ... Finally, the filter is applied to the map, resulting in
panel...

An additional test that can be accomplished with the jackknife map, is
the accuracy of the individual map-pixel variances. If the calculation
is correction, the variance in each jackknife pixel,
$\sigma^2_\mathrm{j}$, is simply the sum of the variances in the even
and odd half-maps, $\sigma^2_\mathrm{e} + \sigma^2_\mathrm{o}$. If a
S/N map is made from the jackknife (i.e., normalizing by
$\sigma_\mathrm{j}$, the expected distribution is simply a Gaussian
with mean zero, and unit standard deviation. A histogram of the
central region of this map is shown in Fig. ???

... fig. showing S/N histogram of the jackknife...

Since the

Both map-making itself, and the whitening filter, will suppress (and
add sidelobes), to the nominal PSF for a point source.
